---
title: "R Notebook"
output: html_notebook
---

```{r, include=FALSE}
library(ape)
library(phytools)
library(dplyr)
library(tibble)
library(purrr)
library(taxize)
```

##### CUSTOM TIP PRUNING #####
This section prunes tips based on a custom set of tip names. This is primarily
to remove duplicates or to remove tips that should not be part of the final
set.

Read in the reference trees to do custom tip pruning
```{r}
get_tips <- function(fin, fout) {
  ##########
  ## PURPOSE:
  #     Read in a reference tree and get the tips of that tree
  #
  ## USAGE:
  #     get_tips(fin, fout)
  #       fin: Name of tree file
  #       fout: Name of output file
  ##########
  
  tree <- read.tree(fin)
  tip_labels <- tree$tip.label
  df <- data.frame(tips = tip_labels)
  
  # Comment this out if you would do not want a list of tips in the tree
  write.table(df, file = fout, sep = "\n", quote = FALSE, row.names = FALSE, col.names = FALSE)
  
  return(tree)
}

jgi <- get_tips("unmodified_references/JGI_2024_unedited.nwk", "JGI_tips.csv")
li <- get_tips("unmodified_references/Lietal_2021_unedited.treefile", "Lietal_2021_tips.csv")
looney <- get_tips("unmodified_references/Looneyetal_2018.nwk", "looney_tips.csv")

russulaceae <- get_tips("Looneyetal_2018.nwk", "looney_tips.csv")
lactarius <- get_tips("lactarius_its2.treefile", "lactarius_tips.csv")
lactifluus <- get_tips("lactifluus_its2.treefile", "lactifluus_tips.csv")
multifurca <- get_tips("multifurca_its2.treefile", "multifurca_tips.csv")
russula <- get_tips("russula_its2.treefile", "russula_tips.csv")
```

Set tips to delete
```{r}
jgi_tips <- c("Acain1", "Acema1", "Amphiamblys_sp.", "Armga1", "Armost1", "Arnium", "Arxula", "Ascni1", "Aspacu1", "Aspbru1", "Aspcos1", "Aspell1", "Aspergillus_nomius3137_1", "Aspeuc1", "Aspfij1", "Asphet1", "Asphom1", "Aspibe1", "Aspind1", "Aspjap1", "Asplac1", "Aspneo1", "Asppip1", "Asppseute1", "Aspsac1", "Aspscl1", "Aspscle1", "Aspuva1", "Aspvad1", "Aspvio1", "Bimnz1", "Biscogniauxia_sp.", "Cadophora_sp.", "Cadophora_sp.", "Canmem1", "CerAGI", "Cersp1", "Chaetomium_sp.", "Chytriomyces_sp.", "Cladorrhinum_sp.", "Cladorrhinum_sp.", "Coniochaeta_sp.", "Copmar1", "Copmic2", "Coprinopsis_sp.", "Copromyces_sp.", "Cordyceps_sp.", "Crycu1", "Cryte1", "Daldinia_sp.", "Decga1", "Densp1", "Diohu1", "Emericellopsis_sp.", "Endogone_sp", "Erebi1", "Fennellomyces_sp.", "FoxFo5176", "FoxII5", "Fusarium_oxysporum_2", "Fusarium_sp.", "Ganoderma_sp.", "Haplosporangium_sp.", "Haplosporangium_sp.", "Haplosporangium_sp.", "Helsul1", "Hydru2", "Hypoxylon_sp.", "Hypoxylon_sp.", "Hypoxylon_sp.", "Hypoxylon_sp.", "Hypoxylon_sp.", "Hypoxylon_sp.", "Hypoxylon_sp.", "Hypoxylon_sp.", "Hypoxylon_sp.", "Hypoxylon_sp.", "Ilyonectria_sp.", "Ilyonectria_sp.", "Jamsp1", "Lachancea_sp.", "Lacqui1", "Lactifluus_cf.", "Lactifluus_cf.", "Leisp1", "Leotiomycetes_sp.", "Leotiomycetes_sp.", "Leptodontidium_sp.", "Linin1", "Lizem1", "Marfi1", "Maseb1", "Meimi1", "MicG_I_3", "Micmi1", "Morhum1", "Morimp1", "Mortierella_sp.", "Mortierella_sp.", "Mortierella_sp.", "Mortierella_sp.", "Mortierella_sp.", "Mortierella_sp.", "Mortierella_sp.", "Mortierella_sp.", "Mortierella_sp.", "Mortierella_sp.", "Mortierella_sp.", "Mortierella_sp.", "Mortierella_sp.", "Mortierella_sp.", "Mortierella_sp.", "Mycgal1", "Mycosphaerella", "Mytre1", "Nemania_sp.", "Nemania_sp.", "Nemania_sp.", "Orpinomyces_sp.", "Panru1", "Paxam1", "Pecoramyces_sp.", "Peniophora_sp.", "Perma1", "Pestalotiopsis_sp.", "Phlebopus_sp.", "Phyllosticta_sp.", "Piromyces_sp.", "Pisolithus_sp.", "Pleery1", "Plucer1", "Podospora_aff.", "Polbr1", "Pseudogymnoascus_sp.", "Pseudogymnoascus_sp.", "Pseudogymnoascus_sp.", "Pseudogymnoascus_sp.", "Pseudogymnoascus_sp.", "Pseudogymnoascus_sp.", "Pyrenochaeta_sp.", "Rachicladosporium_sp.", "Rhexocercosporidium_sp.", "Rhihy1", "Rhiir2_1", "RhiirA1_1", "RhiirB3", "RhiirC2", "Rhili1", "Rhodotorula_sp.", "Rhodotorula_sp.", "Rhodsp1", "Ricmel1", "Saccharomycetaceae_sp", "Setho1", "Sistotrema_sp.", "Sorhu1", "Stagonospora_sp.", "Suidec1", "Suigr1", "Suihi1", "Suillus_cf.", "Synps1", "Terbo2", "Termitomyces_sp.", "Tescy1", "Thasp1", "Thega1", "Tilwa1", "Trace1", "Traci1", "Tribi1", "Tubbor1", "Umbelopsis_sp.", "Umbra1", "Ustsp1", "Xerba1", "Xylaria_cf.", "Xylaria_cf.", "Xylaria_sp.", "Xylaria_sp.", "Xylaria_sp.", "Xylaria_sp.", "Xylaria_sp.", "Xylaria_sp.", "Xylariaceae_sp.", "Xylariaceae_sp.", "Xylariaceae_sp.", "Xylariaceae_sp.", "Xylariaceae_sp.", "Xylariaceae_sp.", "Xylariaceae_sp.", "Xylariaceae_sp.", "Xylariaceae_sp.", "Xylariales_sp.", "Xylariomycetidae_sp.", "Xylariomycetidae_sp.", "Yarrowia_lipolytica_1", "Gelatoporia")

li_tips <- c("Gaeumannomyces_sp._JS-464", "Magnaporthe_sp._MG12", "Magnaporthe_sp._MG03", "Magnaporthe_sp._MG07", "Magnaporthe_sp._MG05", "Magnaporthe_sp._MG08", "Raffaelea_sp._RL272", "Juglanconis_sp._DMW523", "Nectria_sp._B-13", "Fusarium_sp._FIESC_12", "Fusarium_sp._FIESC_28", "Fusarium_sp._JS626", "Fusarium_sp._Na10", "Fusarium_sp._FSSC_6", "Fusarium_sp._JS1030", "Epichloe_sp._AL9924", "Tolypocladium_sp._Salcha_MEA-2", "Tolypocladium_sp._Sup5_PDA-1", "Ophiocordyceps_sp._camponoti-leonardi", "Ophiocordyceps_sp._camponoti-saundersi", "Cordyceps_sp._RAO-2017", "Lecanicillium_sp._MT-2017a", "Lecanicillium_sp._LEC01", "Escovopsis_sp._AC", "Escovopsis_sp._Ae724", "Escovopsis_sp._Ae720", "Escovopsis_sp._Ae733", "Escovopsis_sp._TC", "Colletotrichum_sp._JS-367", "Xylaria_sp._MSU_SB201401", "Xylaria_sp._JS573", "Hypoxylon_sp._EC38", "Hypoxylon_sp._CO27-5", "Hypoxylon_sp._E7406B", "Daldinia_sp._EC12", "Hypoxylon_sp._CI-4A", "Pestalotiopsis_sp._JCM_9685", "Cadophora_sp._DSE1049", "Helotiales_sp._F229", "Phialophora_cf._hyalina_BP_5553", "Rutstroemia_sp._NJR-2017a_BBW", "Rutstroemia_sp._NJR-2017a_WRK4", "Rutstroemia_sp._NJR-2017a_BVV2", "Pseudogymnoascus_sp._VKM_F-4515_FW-2607", "Pseudogymnoascus_sp._VKM_F-4517_FW-2822", "Pseudogymnoascus_sp._BL308", "Pseudogymnoascus_sp._23342-1-I1", "Pseudogymnoascus_sp._VKM_F-4520_FW-2644", "Pseudogymnoascus_sp._VKM_F-4518_FW-2643", "Pseudogymnoascus_sp._VKM_F-4281_FW-2241", "Pseudogymnoascus_sp._04NY16", "Pseudogymnoascus_sp._24MN13", "Pseudogymnoascus_sp._VKM_F-103", "Pseudogymnoascus_sp._VKM_F-4519_FW-2642", "Pseudogymnoascus_sp._05NY08", "Pseudogymnoascus_sp._03VT05", "Pseudogymnoascus_sp._WSF_3629", "Pseudogymnoascus_sp._VKM_F-4513_FW-928", "Pseudogymnoascus_sp._VKM_F-4246", "Pseudogymnoascus_sp._BL549", "Pseudogymnoascus_sp._VKM_F-3775", "Pseudogymnoascus_sp._VKM_F-4516_FW-969", "Pseudogymnoascus_sp._VKM_F-4514_FW-929", "Pseudogymnoascus_sp._VKM_F-3557", "Pseudogymnoascus_sp._VKM_F-3808", "Wickerhamomyces_sp._YB_2243", "Crepidotus_sp._BD2015", "Tricholoma_sp._MG99", "Tricholoma_sp._MG77", "Termitomyces_sp._MG16", "Termitomyces_sp._MG148", "Termitomyces_sp._J132", "Termitomyces_sp._MG145", "Termitomyces_sp._JCM_13351", "Collybia_sp._MG36", "Leucoagaricus_sp._SymC.cos", "Agaricales_sp._HM26F1", "Fibularhizoctonia_sp._CBS_109695", "Boletus_sp._MG95", "Boletus_sp._MG55", "Suillus_sp._MG131", "Rhizopogon_sp._AM2018a", "Ganoderma_sp._BRIUMSc", "Trametes_sp._AH282", "Sarcodon_sp._MG97", "Lactarius_sp._MG50", "Russula_sp._MG48", "Lactarius_sp._MG121", "Peniophora_sp._CONT", "Peniophora_sp._CBMAI_1063", "Dentipellis_sp._KUC8613", "Albatrellus_sp._MG142", "Ramaria_sp._MG151", "Gomphus_sp._MG54", "Ramaria_cf._rubripermanens", "Cryptococcus_cf._gattii", "Cryptococcus_sp._DSM_27421", "Cryptococcus_sp._JCM_24511", "Pseudozyma_sp._F8B2", "Moesziomyces_sp._F16C1", "Moesziomyces_sp._F5C1", "Malassezia_sp.", "Moniliella_sp._wahieum", "Rhodotorula_sp._CCFEE_5036", "Rhodotorula_sp._JG1b", "Rhodotorula_sp._FNED722", "Gongronella_sp._w5", "Endogone_sp._FLASF59071", "Rhizophagus_sp._MUCL_43196", "Zoopage_sp._CTAll", "Ichthyosporea_sp._XGB_2017a", "Nematocida_sp._ERTm5", "Nematocida_sp._1", "Amphiamblys_sp._WSBS2006", "Spizellomyces_sp._palustris", "Piromyces_sp._E2", "Curvularia_sp._IFB-Z10", "Pyrenochaeta_sp._DS3sAY3a", "Pyrenochaeta_sp._UM_256", "Shiraia_sp._slf14", "Paraphoma_sp._B47-9", "Pleosporales_sp._UM_1110_2012", "Stagonospora_sp._SRC1lsM3a", "Rachicladosporium_sp._CCFEE_5018", "Cladosporium_sp._SL-16", "Preussia_sp._BSL10", "Mycosphaerella_sp._PB-2012b", "Mycosphaerella_sp._Ston1", "Cercospora_cf._sigesbeckiae", "Aureobasidium_sp._FSWF8-4", "Eurotiomycetes_sp.", "Herpotrichiellaceae_sp._UM238", "Blastomyces_sp._MA-2018", "Emmonsia_sp._CAC-2015a", "Byssochlamys_sp._BYSS01", "Aspergillus_sp._Z5", "Aspergillus_sp._MA_6037", "Eurotiomycetes_sp._MA_6039", "Eurotiomycetes_sp._MA_6038", "Penicillium_sp._MA_6040", "Penicillium_sp._HKF2", "Penicillium_sp._CF05", "Penicillium_sp._MA_6036", "Byssochlamys_sp._AF001")
```

Delete tips
```{r}
prune_tips <- function(tree, tips, fout) {
  new_tree <- drop.tip(tree, tips)
  write.tree(new_tree, fout)
}

prune_tips(jgi, jgi_tips, "jgi_half_pruned.txt")
prune_tips(li, li_tips, "Lietal_2021.nwk")
```
##############################

##### TRIM TO FAMILY LEVEL #####
This section creates a family-level backbone. For other levels of taxonomy, replace
the tips with sets of tip names based on the level you need.

1. Get family information of tips in reference trees
```{r}
get_fams <- function(trefile) {
  # read in tree file to trim
  tre <- read.tree(trefile)
  
  # get taxonomy information from UNITE database
  all_tax <- read.csv("database_information/unite_taxonomy.csv")
  all_tax <- distinct(all_tax, Species, .keep_all = TRUE)
  
  # manipulate taxonomy information to get genera
  tre_tips <- as.data.frame(tre$tip.label)
  tre_tips[, 'empty'] = NA
  colnames(tre_tips) <- c("Species", "Genus")
  tre_tips$Genus <- sub("_.*", "", tre_tips$Species)

  tre_tax <- tre_tips %>%
    inner_join(all_tax, by=c("Genus"))

  # extract family information using genera
  tre_tax <- subset(tre_tax, select = c(Species.x, Family))
  tre_tax <- distinct(tre_tax, Species.x, .keep_all = TRUE)
  colnames(tre_tax) <- c("Species", "Family")

  # species in tre that do not have family information in the UNITE database
  # after being genus-matched
  non_unite <- subset(tre_tips, !(Species %in% tre_tax$Species))
  
  # return taxa that have family information and those that don't
  return(list("tre_tax" = tre_tax, "non_unite" = non_unite))
}

jgi_tax <- get_fams("JGI.nwk")
li_tax <- get_fams("Lietal_2021.nwk")
```

2. Get missing family information from NCBI
```{r}
get_ncbi <- function(input_sp) {
  tre_sp <- input_sp$non_unite$Species
  tre_ids <- get_ids(tre_sp, db = "ncbi")
  fam <- classification(tre_ids, db = "ncbi")
  return(fam)
}

jgi_missing_fams <- get_ncbi(jgi_tax)
# print the taxonomy information for missing families in JGI tree
for(i in 1:length(jgi_missing_fams$ncbi)) {
  print(jgi_missing_fams$ncbi[i])
}

li_missing_fams <- get_ncbi(li_tax)
# print the taxonomy information for missing families in Li tree
for(i in 1:length(li_missing_fams$ncbi)) {
  print(li_missing_fams$ncbi[i])
}

# from here, manually add taxonomy information to unite_taxonomy.csv and rerun
# previous cell (step 1)
```

3. get backbone for tree
```{r}
### genus-level backbone for Russulaceae family tree
# Steps:
# 1. Keep all tips except 1 per genus
# 2. Save as backbone tree
# 3. Rename the 1 species per genus to "genus_genusName"
#
# Notes:
# - the resulting backbone will not have accurate branch lengths, but will have accurate topology
# - R can only handle a certain # of characters per line. Drop ~100 tips at a time per line to bypass this limit.

russulaceae <- drop.tip(russulaceae, c("insert all species here"))

write.tree(russ_tree, file="russulaceae_backbone.nwk")

```

################################